---
title: "Assingment 11"
author: "ABDUL RAUF"
date: "2023-01-19"
output: html_document
---

\begin{center} * LAB - 11 * \end{center}
\newpage

                        __NAME__:-**ABDUL RAUF**
                        __ENRL NO__:-**GL6092** 
                        __ROLL NO__:-**22DSMSA116**    




# Assingment 1

## Read the data from excel


```{r}
advert<-read.csv("Advertising.csv")
head(advert)
names(advert)
attach(advert)
pairs(advert)
dim(advert)
```

## Fitting the model 

```{r}
m1<-lm(sales~TV+radio+newspaper,data = advert)
m1
summary(m1)
AIC(m1)
```

$\bullet$ Except Newspaper all the variables are significant as their p- values are less than **0.05** .

$\bullet$ 89.56% of the variability of sales is explained by the model.



```{r}
m2<-lm(sales~TV+radio,data = advert)
m2
summary(m2)
AIC(m2)
```

$\bullet$ 89.62% of the variability of sales is explained by the model.

$\bullet$ m2 has the lowest AIC value and in this model all the variables are significant.

$\bullet$ adjusted R sq and MSE of m2 have approximately same value as compared to the m1 so m2 is better as compared to model m1.

## Bootstraping

```{r}
library(boot)
set.seed(1)
advert.fn<-function(advert,index)
  coef(lm(sales~TV+radio,data=advert,subset=index))
advrt.obj<-boot(data = advert,statistic = advert.fn ,R=200)
advrt.obj

```
Estimates of parameters are as follows:-

$\bullet$ $SE(\hat{\beta_0})$ = **0.3133033**

$\bullet$ $SE(\hat{\beta_1})$ = **0.002003**

$\bullet$ $SE(\hat{\beta_2})$ = **0.010484**

Estimates of the bais of the parameter are as follows:-

$\bullet$ $Bais(\hat{\beta_0})$ = **-4.812417e-03**

$\bullet$ $Bais(\hat{\beta_1})$ = **7.985856e-05 **

$\bullet$ $Bais(\hat{\beta_2})$ = **-9.421676e-05**



# Assingment 2

## Read the data from excel

```{r}
sol<-read.csv("solubility.csv")
head(sol)
names(sol)
attach(sol)
pairs(sol)
dim(sol)
```

## Fitting the model

```{r}
M1<-lm(y~X1+X2+X3,data = sol)
M1
summary(M1)
AIC(M1)
``` 
$\bullet$ 84.86% of the variability of y is explained by the model. and all the variables are significant.

## Bootstraping

```{r}
library(boot)
set.seed(1)
sol.fn<-function(sol,index)
  coef(lm(y~X1+X2+X3,data=sol,subset=index))
sol.obj<-boot(data = sol,statistic = sol.fn ,R=200)
sol.obj

```


Estimates of parameters are as follows:-

$\bullet$ $SE(\hat{\beta_0})$ = **0.187210926**

$\bullet$ $SE(\hat{\beta_1})$ = **0.018998236**

$\bullet$ $SE(\hat{\beta_2})$ = **0.010711207**

$\bullet$ $SE(\hat{\beta_3})$ = **0.006702752**


Estimates of the bais of the parameter are as follows:-

$\bullet$ $Bais(\hat{\beta_0})$ = **0.0401583977**

$\bullet$ $Bais(\hat{\beta_1})$ = **-0.0034272314 **

$\bullet$ $Bais(\hat{\beta_2})$ = **-0.0005539604**

$\bullet$ $Bais(\hat{\beta_2})$ = **-0.0022745225**





