---
title: "lab_11-BOOTSTRAP"
author: "Vaishnavi singh"
date: "JAN 11,23"
output:
  pdf_document: default
  html_document: default
---
\pagenumbering{roman}
\newpage
\pagenumbering{arabic}



# ASSIGNMENT 1 

## Reading the data
```{r}
Data<-read.csv("Advertising.csv")
head(Data)
names(Data)

```
## Fitting the MLR model
```{r}
M1<-lm(sales~TV+radio+newspaper,data=Data)
summary(M1)
```
For this model,model adequacy is 89.56% but two variables are non significant.Thus,this model is not good fit.
```{r}
M2<-lm(sales~TV+radio,data=Data)
summary(M2)
```
Since for this model,coefficient of determination is 89.62% with all variables to be significant.Thus,this model is good fit. 



```{r}
library(boot)
```

```{r}
boot.fn<-function(Data,index)
  coef(lm(sales~TV+radio,data=Data,subset=index))
boot.fn(Data,sample(200,200,replace=TRUE))

set.seed(19)
boot(Data,boot.fn,R=200)

```
This indicates that the bootstrap estimates for $$SE(\hat\beta_0)=0.3255$$,
$$SE(\hat\beta_1)=0.0019$$ and $$SE(\hat\beta_2)=0.0098$$
\newpage

# ASSIGNMENT 2 

## Reading the data
```{r}
Data<-read.csv("Datachemical.csv")
names(Data)
head(Data)
```

## Fitting the model
```{r}
M1<-lm(y~.,data=Data)
summary(M1)
```

This model explains only 47.71% variability and two variables are non significant for the model.Hence,this model is not good fit.
```{r}
M2<-lm(y~acidconc+acidtemp+sulfideconc,data=Data)
summary(M2)
```
This model is better as it explains 46.72% variability with all variables significant.

```{r}
boot.fn<-function(Data,index)
  coef(lm(y~acidconc+acidtemp+sulfideconc,data=Data,subset=index))
boot.fn(Data,sample(200,200,replace=TRUE))

set.seed(19)
boot(Data,boot.fn,R=200)

```
This indicates that the bootstrap estimates for 
$$SE(\hat\beta_0)=10.96$$,
$$SE(\hat\beta_1)=8.2514$$,
$$SE(\hat\beta_2)=0.2082$$ and
$$SE(\hat\beta_3)=30.9324$$ .

