---
title: "MULTIPLE LINEAR REFRESSION MODEL"
author: "Mohammed Swaned M"
date: "05-12-2022"
output:
  pdf_document: default
  word_document: default
  html_document: default
---
\tableofcontents
\newpage


Name         : Mohammed Swaned M

Enrollment_no: GN4994

Faculty_no   : 22DSMSA104





# Multiple Linear Regression Model
## Introduction

The Multiple linear regression model is defined as
$y=\beta_0+\beta_1 \times X_1 +\cdot\cdot\cdot + \beta_pX_p +\epsilon$ 

Now we consider the wage12 data and the multiple linear regression model for that data is given as

$wage=\beta_0+\beta_1\times education +\beta_2\times tenure +\beta_3\times nonwhite + \beta_4\times female +\beta_5 \times married +\epsilon$

Now read data

```{r warning=FALSE}
wage12<-wooldridge::wage1[c(1,2,3,4,5,6,7)]
head(wage12)
```
## Make the scatter plot
```{r warning=FALSE}
pairs(wage12)
```
From the above plot, we may see that there is a linear relationship between TV and sales, radio``andsales’ and newspaper and sales. To figure out more we obtain the correlation coefficient among the
variables.

## Calculate the correlation coefficient
```{r warning=FALSE}
cor(wage12)
```
## Fitting of multiple Linear Regression Model
To estimate the coefficients of the variables educ,exper,tenure,nonwhite,female and married, we fit the following model
```{r warning=FALSE}
M1=lm(wage~educ+exper+tenure+nonwhite+female+married, data = wage12)
summary(M1)
```
## Checking Significance of Model
To check the significance of the model, we check F statistic and for that we set the hypotheses as follows:
Null Hypothesis: $H_0 : \beta_1 = \beta_2 =\cdot\cdot\cdot = \beta_k = 0$

Alternative Hypothesis: $H_1 : At least \:one \:\beta_i \neq 0, i = 1, 2, \cdot\cdot\cdot , k$.
F-statistic: 50.41 on 6 and 519 DF, p-value: $< 2.2e-16$ Since the F statistic is 50.41 on 6 and 519 df with p-value < 2.2e-16 i.e. almost zero. Hence we *reject the null hypothesis* that means there is at least one$\beta_i$ that is not equal to zero. Therefore, our model is significant.

## Checking Significance of Variables
To check the significance of variable(s), we check the $t-ratios$ and its corresponding $p-values$. We set the hypotheses as follows:
$H_0 : \beta_i = 0\:\: Vs \space H_1 : \beta_i \neq 0\: where \:i = 0, 1, 2, 3$ 
Now we check the t-statistics and p value of the corresponding variables one by one and decide that which
one is significant. SO the p-value for *intercept* term is .0289 that is almost zero, hence we reject the null hypothesis and accept that $\beta_0 \neq 0$.
The p-value of *Education* is  $2\times10^{-16}$that is almost zero, hence we reject the null hypothesis and accept that $\beta_1 \neq 0$
The p-value of *Tenure* is $1.29\times10^{-10}$ that is almost zero, hence we reject the null hypothesis and accept that $\beta_3 \neq 0$.
The p-value of *Female* is $1.57\times10^{-10}$ that is almost zero, hence we reject the null hypothesis and accept that $\beta_5 \neq 0$.
The p-value of *Experience* is 0.1201 that is greater than 0.05, hence we fail to reject the null hypothesis and accept that $\beta_3 = 0$.
The p-value of *Nonwhite* is 0.8775 that is greater than 0.05, hence we fail to reject the null hypothesis and accept that $\beta_2 = 0$.
The p-value of *married* is 0.0525 that is greater than 0.05, hence we fail to reject the null hypothesis and accept that $\beta_6 = 0$.
This means that the variables *Experience,Nonwhite and Married*is not significant in this model.
Since these variables is not significant, so we remove these  variable from the model.
Hence our final model is:
$wage= \beta_0 + \beta_1 \times education + \beta_3\times tenure + \beta_5 \times female + \epsilon$
```{r warning=FALSE}
M2=lm(wage~educ+tenure+female,data=wage12)
summary(M2)
```
Here,The p-value of *intercept* is 0.193  that is greater than 0.05, hence we fail to reject the null h*ypothesis and accept that $\beta_0= 0$.
since *intercept* is not significant,so we remove the variable from the model
Hence our final model is:
$wage=  \beta_1 \times education + \beta_3\times tenure + \beta_5 \times female + \epsilon$

```{r warning=FALSE}
M3=lm(wage~educ+tenure+female-1,data=wage12)
summary(M3)
```
## Adequacy of Model
### Residual Standerd Error

$RSE=\sqrt(\frac{RSS}{n-p-1})$

where $RSS=\sqrt(\sum_{i=1}^{n}(y_i -\hat y_i)^2$ and is called residual sum of squares(RSS)
```{r warning=FALSE}
wage=wage12$wage
educ=wage12$educ
tenure=wage12$tenure
female=wage12$female
wagehat=M3$coefficients[1]*educ+M3$coefficients[2]*tenure+M3$coefficients[3]*female
#alternaively
wagehat2=M3$fitted.values
residManual=wage-wagehat
rssMan=sum(residManual^2)
rssMan  
n=length(educ)
p=3
rseman=sqrt(rssMan/n-p-1)
rseman
resid=M3$residuals
rss=sum(resid^2)
rss
n=length(wage12$educ)
p=3
rse=sqrt(rss/n-p-1)
rse
```
Now Calculate RSE from the rectified model

```{r warning=FALSE}
resid2=M3$residuals
rss2=sum(resid2^2)
rss2
p=3
rse2=sqrt(rss/n-p-1)
rse2
```
### R Squared

Multiple R-squared: 0.3682, Adjusted R-squared: 0.3609. The reported R squared is 0.3682 that is approximately 0.37. So we see that 37% variability of wage is explained by education, experience, tenure,nonwhite,female and married and the adjusted
R- squared is 0.8956 when insignificant variables (experience,nonwhite and married) is attached.
After omitting these insignificant variables from the model and we examine the R-squared and
adjusted R squared. They are as follows:
Multiple R-squared: 0.3577, Adjusted R-squared: 0.3544.Here the intercept is not a significant variable .After ommitting the insignificant variables from the variables and we examine the R-squared and adjusted R-squared.They are as follows:
Multiple R-squared: 0.8187, Adjusted R-squared: 0.8176.We find that there is no difference in R squared but adjusted R-squared has increased a little bit. That is theevidence that if we remove any insignificant variable from the model, them adjusted R-squared increased.

## Dummy Variable
### Female as dummy variable
 Here we model the data as follows: 
 $Wage=\beta_0+\beta_1\times female$
 
```{r warning=FALSE}
M4=lm(wage~female,data = wage12)
summary(M4)
```
Table displays the coefficient estimates and other information associated with the model. So the  Model of the data as follows:
$Wage=7.0995-2.5118\times female$
However, we notice that the p-value for the dummy variable that is $1.042\times 10^{-15}$. This means that the variable *female* is significant.

### Nonwhite as dummy variable
Here we model the data as follows: 
 $Wage=\beta_0+\beta_1\times nonwhite$
 
```{r warning=FALSE}
M5=lm(wage~nonwhite,data = wage12)
summary(M5)
```
Table displays the coefficient estimates and other information associated with the model. So the  Model of the data as follows:
$Wage=5.9442-0.4682\times nonwhite$
However, we notice that the p-value for the dummy variable is high that is $0.378$ This means that the variable *nonwhite* is insignificant.
### Married as dummy variable
Here we model the data as follows: 
 $Wage=\beta_0+\beta_1\times married$
 
```{r warning=FALSE}
M6=lm(wage~married,data = wage12)
summary(M6)
```
Table displays the coefficient estimates and other information associated with the model. So the  Model of the data as follows:
$Wage=4.8439+1.7296\times married$
However, we notice that the p-value for the dummy variable is $1.12\times10^{-7}$ This means that the variable *married* is significant.