---
title: "MULYIPLE LINEAR REGRESSION MODEL"
author: "Mohammed Swaned M"
date: "2022-12-05"
output:
  pdf_document: default
  html_document: default
---

Name         : Mohammed Swaned M

Enrollment_no: GN4994

Faculty_no   : 22DSMSA104





# Multiple Linear Regression Model
## Introduction

The Multiple linear regression model is defined as
$y=\beta_0+\beta_1 \times X_1 +\cdot\cdot\cdot + \beta_pX_p +\epsilon$ 

Now we consider the wage12 data and the multiple linear regression model for that data is given as

$wage=\beta_0+\beta_1\times education +\beta_2\times tenure +\beta_3\times nonwhite + \beta_4\times female +\beta_5 \times married +\epsilon$

Now read data

```{r}
credit=read.csv("credit.csv")
head(credit)
credit1=credit[c(-7,-8,-9,-10)]
credit1
```
## Make the scatter plot
```{r}
pairs(credit1)
```
From the above plot, we may see that there is a linear relationship between TV and sales, radio``andsales’ and newspaper and sales. To figure out more we obtain the correlation coefficient among the
variables.

## Calculate the correlation coefficient
```{r}
cor(credit1)
```
## Fitting of multiple Linear Regression Model
To estimate the coefficients of the variables TV, radio and newspaper, we fit the following model
```{r}
M1=lm(Balance~., data = credit)
summary(M1)
```
## Checking Significance of Model
To check the significance of the model, we check F statistic and for that we set the hypotheses as follows:

Null Hypothesis:$H_0 : \beta_1 = \beta_2 =\cdot\cdot\cdot = \beta_k = 0$

Alternative Hypothesis:$H_1 : At least \:one \:\beta_i \neq 0, i = 1, 2, \cdot\cdot\cdot , k$.

F-statistic: 750.3 on 11 and 388 DF, p-value: $< 2.2e-16$ Since the F statistic is 750.3 on 11 and 388 df with p-value < 2.2e-16 i.e. almost zero. Hence we *reject the null hypothesis* that means there is at least one$\beta_i$ that is not equal to zero. Therefore, our model is significant.

## Checking Significance of Variables
To check the significance of variable(s), we check the $t-ratios$ and its corresponding $p-values$. We set the hypotheses as follows:
$H_0 : \beta_i = 0\:\: Vs \ H_1 : \beta_i \neq 0\: where \:i = 0, 1, 2, 3$ 
Now we check the t-statistics and p value of the corresponding variables one by one and decide that which
one is significant. SO the p-value for *intercept* term is $2\times10^{-16}$ that is almost zero, hence we reject the null hypothesis and accept that $\beta_0 \neq 0$.
The p-value of *Income* is  $2\times10^{-16}$that is almost zero, hence we reject the null hypothesis and accept that $\beta_1 \neq 0$
The p-value of *Limit* is $1.21\times10^{-8}$ that is almost zero, hence we reject the null hypothesis and accept that $\beta_3 \neq 0$.
The p-value of *Rating* is $0.0211$ that is  less than 0.05, hence we reject the null hypothesis and accept that $\beta_5 \neq 0$.
The p-value of *Cards* is $5.40\times10^{-5}$ that is almost zero, hence we reject the null hypothesis and accept that $\beta_5 \neq 0$.
The p-value of *Age* is $0.034$ that is less than 0.05, hence we reject the null hypothesis and accept that $\beta_5 \neq 0$.
The p-value of *Student* is $2\times10^{-16}$ that is almost zero, hence we reject the null hypothesis and accept that $\beta_5 \neq 0$.
The p-value of *Education* is 0.4921 that is greater than 0.05, hence we fail to reject the null hypothesis and accept that $\beta_3 = 0$.
The p-value of *Own* is 0.2832 that is greater than 0.05, hence we fail to reject the null hypothesis and accept that $\beta_2 = 0$.
The p-value of *Married* is 0.4107 that is greater than 0.05, hence we fail to reject the null hypothesis and accept that $\beta_6 = 0$.
The p-value of *Region* is 0.4083 that is greater than 0.05, hence we fail to reject the null hypothesis and accept that $\beta_6 = 0$.
This means that the variables *Education,Own,Married and Region*is not significant in this model.
Since these variables is not significant, so we remove these  variable from the model.
Hence our final model is:
$Balance = \beta_0 + \beta_1 \times Income+ \beta_2\times Limit + \beta_3 \times Ratings + \beta_4 \times Cards + \beta_5 \times Age + \beta_8 \times Student + \epsilon$

```{r}
M2=lm(Balance~Income+Limit+Rating+Cards+Age+Student,data=credit)
summary(M2)
```
## Adequacy of Model
### Residual Standerd Error

$RSE=\sqrt(\frac{RSS}{n-p-1})$

where $RSS=\sqrt(\sum_{i=1}^{n}(y_i -\hat y_i)^2$ and is called residual sum of squares(RSS)
```{r}
Balance=credit$Balance
Income=credit$Income
Limit=credit$Limit
Rating=credit$Rating
Cards=credit$Cards
Age=credit$Age
Balancehat=M2$coefficients[1]*Income+M2$coefficients[2]*Limit+M2$coefficients[3]*Rating+M2$coefficients[4]*Cards+M2$coefficients[5]*Age
#alternaively
resid=M2$residuals
rss=sum(resid^2)
rss
n=length(credit$Income)
p=5
rse=sqrt(rss/n-p-1)
rse
```
Now Calculate RSE from the rectified model

```{r}
resid2=M2$residuals
rss2=sum(resid2^2)
rss2
p=5
rse2=sqrt(rss/n-p-1)
rse2
```
### R Squared

Multiple R-squared: 0.9551, Adjusted R-squared: 0.9538. The reported R squared is 0.9551 that is approximately 0.95. So we see that 95% variability of Balance is explained by Income, Limit, Rating, Cards, Age,Education,Own,Student,Married and Region and the adjusted
R- squared is 0.9538 when insignificant variables(Education,Own,Married and Region ) is attached.
After omitting these insignificant variables from the model and we examine the R-squared and adjusted R squared. They are as follows:
Multiple R-squared: 0.9547, Adjusted R-squared: 0.954.We find that there is no difference in R squared but adjusted R-squared has increased a little bit. That is theevidence that if we remove any insignificant variable from the model, them adjusted R-squared increased.

## Dummy Variable
### Studnt as dummy variable
 Here we model the data as follows: 
 $Balance =\beta_0+\beta_1\times Student$
 
```{r}
M3=lm(Balance~Student,data = credit)
summary(M3)
```
Table displays the coefficient estimates and other information associated with the model. So the  Model of the data as follows:
$Balance=480.37+396.46 \times Student$

However, we notice that the p-value for the dummy variable that is $1.488\times 10^{-7}$. This means that the variable *Student* is significant.

### Married as dummy variable
Here we model the data as follows: 
 $Balance=\beta_0+\beta_1\times Married$
 
```{r}
M4=lm(Balance~Married,data = credit)
summary(M4)
```
Table displays the coefficient estimates and other information associated with the model. So the  Model of the data as follows:
$Balance=523.290-5.347\times Married$
However, we notice that the p-value for the dummy variable is high that is $0.9099$ This means that the variable *Married* is insignificant.
### Own as dummy variable
Here we model the data as follows: 
 $Balance=\beta_0+\beta_1\times Own$
 
```{r}
M5=lm(Balance~Own,data = credit)
summary(M5)
```
Table displays the coefficient estimates and other information associated with the model. So the  Model of the data as follows:
$Balance=509.80+19.73\times Own$
However, we notice that the p-value for the dummy variable is high that is$0.6685$ This means that the variable *Own* is insignificant.

### Region as dummy variable
Here we model the data as follows: 
 $Balance=\beta_0+\beta_1\times Region$
 
```{r}
M6=lm(Balance~Region,data = credit)
summary(M6)
```
Table displays the coefficient estimates and other information associated with the model. So the  Model of the data as follows:
$Balance=531-12.50\times RegionSouth-18.69\times RegionWest$
However, we notice that the p-value for the dummy variable is high that is$0.9575$ This means that the variable *Region* is insignificant.
